Repeatability of Pre-Flashover Fire Patterns on Gypsum Wall Board

### 2015-03-12 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
TIFF images changed to PDF and incorporated into revised doc with full ToC

\chapter{Scope of Study}

@@

\chapter{Technical Approach}

@@

\subsection{Bench Scale Experiments}

Several types of bench scale experiments were conducted to generate data. This data could be used to assist in the analysis of the fire patterns and potentially be used as input for the FDS models.
Critical Heat Flux – Conducted a series of experiments on samples of painted gypsum wallboard to determine the repeatability of a ``critical heat flux'' that results in a clear line of demarcation, a ``burn/no burn'' line.  The flooring radiant panel (ASTM~E648) was used to examine this value.

@@

\subsubsection{Fire Pattern Recreation}

Heskestad developed an empirical method for calculating the mean visible flame height for fuels which do not have substantial in-depth combustion~\cite{Beyler:1986,Heskestad:SFPE}. This correlation was compared with the flame height results of the source fires and the instrumented wall experiments.

FDS~\cite{McGrattan:2014} and Smokeview~\cite{Forney:2014} were used to examine the ability to numerically reproduce the source fires and the instrumented wall experiments.  Comparison between the experiments and the model will provide valuable validation data for the model.

@@

\section{Characterizing the Source Fires}

@@

\subsubsection{Fuels}

@@

\subsubsection{Experimental Arrangement}

@@

\chapter{Results}

@@

Examples of a set of TIFF images for each fuel are shown in Figures 16 through 18.   Each figure has 56 sequential images from the video segment.  The images total almost 2~s of actual burn time.  Notice how the flame height cycles up and down in each of the sequences.  These images provide additional information and continuity of the motion of the flame so that the analysis is less likely to miss and extreme (minimum or maximum) position when compared to the still photographs.

From each of the peak heat release rate video clips, the edited TIFF files were loaded into Streams.  For the natural gas and gasoline fueled fires, approximately 900 images were used in each image sequence.  The polyurethane foam experiments had approximately 450 images in each sequence for each experiment.  Once the image sequence was created, a series of filters and amplifiers are applied to the images to eliminate all colors but red, based on a method developed by Goble~\cite{Goble:2007}.  Once all of the images were converted to a pure red ``flame'' which represented the area of visible fire, on a black background, the images were processed within Streams to develop an intensity field based on all of the images being ``overlaid'' on each other.  The end result is a contour plot of the intermittency of the flame ranging from zero to one.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{../Figures/Fig19}\\
  \caption[Comparison images of intensity fields derived from the video images of the natural gas fueled fire]{Comparison images of intensity fields derived from the video images of the natural gas fueled fire using Streams. The images, starting with the upper left, were generated with 5 mm, 10 mm, 20 mm, 40 mm and 80 mm resolutions.}
  \label{Intensity}
\end{figure}

@@

\chapter{Instrumented Calibration (non-combustible) Wall Experiments}

@@

\subsection{Heat Flux Wall Temperatures}

\chapter{Instrumented Gypsum Board Wall Experiments}

@@

@@

>>>>

\chapter{Scope of Study}
\label{chapter:Scope of Study}

@@

\chapter{Technical Approach}
\label{chapter:Technical Approach}

@@

\section{Bench Scale Experiments}

Several types of bench scale experiments were conducted to generate data. This data could be used to assist in the analysis of the fire patterns and potentially be used as input for the FDS models.

\subsection{Critical Heat Flux}

Conducted a series of experiments on samples of painted gypsum wallboard to determine the repeatability of a ``critical heat flux'' that results in a clear line of demarcation, a ``burn/no burn'' line.  The flooring radiant panel (ASTM~E648) was used to examine this value.

@@

\section{Fire Pattern Recreation}

The ability to recreate or simulate a fire pattern or certain characteristics of a fire pattern was examined with a variety of tools.  The first characteristic examined was the ability to determine based on the calculated heat flux to the wall if a "visible or measurable physical change" would occur to the painted paper surface of the gypsum wallboard.  The next characteristic of the fire pattern that was examined was the height.  The height of the fire pattern was then compared to the predictive results from flame height correlations. Heskestad developed an empirical method for calculating the mean visible flame height for fuels which do not have substantial in-depth combustion~\cite{Beyler:1986,Heskestad:SFPE}. This correlation was compared with the flame height results of the source fires and the instrumented wall experiments.

FDS~\cite{McGrattan:2014} and Smokeview~\cite{Forney:2014} were used to examine the ability to numerically reproduce the source fires and the instrumented wall experiments. These simulation tools provide results in the form of visual field data, in other they provide results in the same format that the investigator would see or document the fire pattern.  Comparison between the experiments and the model may provide valuable validation data for the model.

@@

\chapter{Characterizing the Source Fires}
\label{chapter:Characterizing the Source Fires}

@@

\section{Fuels}

@@

@@

\chapter{Results}
\label{chapter:Results}

@@

Examples of a set of TIFF images for each fuel are shown in Figures .~\ref{flame_height_comp_FHNG80_1} through .~\ref{flame_height_comp_FHPUF_1}.   Each figure has 56 sequential images from the video segment.  The images total almost 2~s of actual burn time.  Notice how the flame height cycles up and down in each of the sequences.  These images provide additional information and continuity of the motion of the flame so that the analysis is less likely to miss and extreme (minimum or maximum) position when compared to the still photographs.
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../Figures/flame height comp FHNG80-1.png}\\
	\caption[Example of sequential video frames sampled from a natural gas fueled experiment]{Example of sequential video frames sampled from a natural gas fueled experiment.  In this set of images, which spans almost two seconds, the visual continuous flame height fluctuates from approximately 0.6 m to 1.1 m.}
	\label{flame_height_comp_FHNG80_1}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../Figures/flame height comp FHGAS-1.png}\\
	\caption[Example of sequential video frames sampled from a gasoline fueled experiment]{Example of sequential video frames sampled from a gasoline fueled experiment.  In this set of images which spans almost two seconds the visual continuous flame height fluctuates from approximately 0.6 m to 1.1 m.}
	\label{flame_height_comp_FHGAS_1}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../Figures/flame height comp FHPUF-1.png}\\
	\caption[Example of sequential video frames sampled from a polyurethane foam fueled experiment]{Example of sequential video frames sampled from a polyurethane foam fueled experiment.  In this set of images which spans almost two seconds the continuous visual flame height fluctuates from 0.4 m to 0.9 m.}
	\label{flame_height_comp_FHPUF_1}
\end{figure}

From each of the peak heat release rate video clips, the edited TIFF files were loaded into Streams~\cite{Nokes:2011}.  For the natural gas and gasoline fueled fires, approximately 900 images were used in each image sequence.  The polyurethane foam experiments had approximately 450 images in each sequence for each experiment.  Once the image sequence was created, a series of filters and amplifiers are applied to the images to eliminate all colors but red, based on a method developed by Goble~\cite{Goble:2007}.  Once all of the images were converted to a pure red ``flame'' which represented the area of visible fire, on a black background, the images were processed within Streams to develop an intensity field based on all of the images being ``overlaid'' on each other.  The end result is a contour plot of the intermittency of the flame ranging from zero to one.

@@

\begin{figure}[p]
	\centering
	\includegraphics[trim=0in 0in 3.5in 0in,clip=true,width=1.7in]{../Figures/FHNG80-1_GS_5mm_10s_color}
	\includegraphics[trim=0in 0in 3.5in 0in,clip=true,width=1.7in]{../Figures/FHNG80-1_GS_10mm_10s_color}\\
	\includegraphics[trim=0in 0in 3.5in 0in,clip=true,width=1.7in]{../Figures/FHNG80-1_GS_20mm_10s_color}
	\includegraphics[trim=0in 0in 3.5in 0in,clip=true,width=1.7in]{../Figures/FHNG80-1_GS_40mm_10s_color}
	\includegraphics[trim=0in 0in 3.5in 0in,clip=true,width=1.7in]{../Figures/FHNG80-1_GS_80mm_10s_color}\\
	 \caption[Comparison images of intensity fields derived from the video images of the natural gas fueled fire]{Comparison images of intensity fields derived from the video images of the natural gas fueled fire using Streams. The images, starting with the upper left, were generated with 5 mm, 10 mm, 20 mm, 40 mm and 80 mm resolutions.}
	 \label{Intensity}
\end{figure}

@@

\chapter{Instrumented Calibration (non-combustible) Wall Experiments}
Purpose

\section{Experimental Arrangement}
A free standing, non-combustible, calcium silicate wall was instrumented to measure the thermal exposure of the wall from the fire in terms of temperature and heat flux.  The fires were initially placed at predetermined, fixed positions away from the wall. The burner was moved toward the wall until the burner was against the wall.

@@

\subsection{Wall Temperatures}

\subsection{Heat Flux}

\subsection{Fire Pattern}

\chapter{Instrumented Gypsum Board Wall Experiments}
Purpose
\section{Experimental Arrangement}
A free standing, non-combustible, steel frame with gypsum wallboard attached to the surface was instrumented to measure the thermal exposure of the wall from the fire in terms of temperature and heat flux.  The fires were initially placed at predetermined, fixed positions away from the wall. The burner was moved toward the wall until the burner was against the wall.  If the results indicated that no burn pattern would result from a given combination of fuel and position, then that fuel and position combination were eliminated from the remainder of the test matrix.

@@

\subsection{Wall Temperatures}

\subsection{Heat Flux}

\subsection{Fire Pattern}

\chapter{Bench Scale Experiments}

@@

\section{Radiant Flooring Panel Experiments}

\section{Radiant Panel Experiments}

\subsection{Ignition Temperature}

\subsection{Heat Release Rate per Unit Area}

\subsection{Critical Heat Flux}

\chapter{Fire Pattern Compartment Experiments}

\section{Wall Construction Type Experiments}
\subsection{Experimental Arrangement}
\subsection{Results}

\section{Wall Position Experiments}
\subsection{Experimental Arrangement}

\subsection{Results}

\subsubsection{Natural Gas}
\subsubsection{Gasoline}
\subsubsection{Polyurethane Foam}

\section{Corner Position Experiments}

\subsection{Experimental Arrangement}

\subsection{Results}

\subsubsection{Natural Gas}
\subsubsection{Gasoline}
\subsubsection{Polyurethane Foam}

\chapter{Summary of Experimental Results}

Needed?

\chapter{Comparison of Algebraic Fire Characterizations with Fire Pattern Results}

\section{Evidence of a Fire Pattern}

\section{Flame Height Correlation}

\subsection{Correlation Descriptions}

\section{Comparison of Results}

\chapter{Comparison of Field Model Simulations with Experimental Results}

\section{Fire Dynamics Simulator and Smokeview}

\subsection{Simulation of Fires}

\subsection{Input Data}

\section{Comparison of FDS with Source Fire Results}

\subsection{Grid Size Analysis}

\subsection{Plume Temperatures}

\subsection{Heat Flux}

\subsection{Flame Height}

\section{Comparison of FDS with Fire Pattern Results}

\chapter{Discussion}

\chapter{Future Research Needs}

\chapter{Conclusions}
